Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-10-04 paper.tex
Night work

    Streaming currents have been heavily investigated as a means of generating electrical energy from pressure gradients.~\cite{Chang2009,Daiguji2006,Daiguji2004b,Davidson2008a,Davidson2008,CherngHon2012,Jiao2014,Lu2006,Olthuis2005,Osterle1964,Pennathur2007,Ren2008a,VanderHeyden2006,Heyden2007,Xie2008,Yang2003}
    Theoretical predictions of the efficiency of standard micro/nano-fluidic channels are 2\% for pure water and 7\% for sodium chloride.~\cite{VanderHeyden2006}
    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10\thinspace--\SI{16}{\micro\metre}.~\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with \SI{6}{\micro\metre} pores.~\cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a \SI{75}{\nano\metre} by \SI{50}{\micro\metre} silica channel.~\cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a \SI{200}{\nano\metre} pore in an alumina membrane.~\cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a \SI{0.5}{\nano\metre} cylindrical pore in polyethylene terephalate foil.~\cite{Xie2008}
    \end{itemize}
    These results indicate that small channels using solutions containing salt are more efficient.
    According to \cite{Daiguji2004b}, the efficiency is maximised when the channel height is about half the Debye length.
    For water with a sodium content in the range 1--100~ppm this equates to a channel height in the order of 100--10~nm.
    Additionally, ~\cite{VanderHeyden2006} states that the maximum efficiency is found when the salt concentration is low.

@@

    \section{Sizing of Streaming Cell Harvester}
    \label{sect:harvesterSize}
    In order to achieve the target conversion efficiency of 1\% the internal channel height must be reduced to something in the order of \SI{1}{\micro\meter}--\SI{10}{\nano\meter}.

>>>>

    Streaming currents have been heavily investigated as a means of generating electrical energy from pressure gradients.~\cite{Chang2009,Daiguji2006,Daiguji2004b,Davidson2008a,Davidson2008,CherngHon2012,Jiao2014,Lu2006,Olthuis2005,Osterle1964,Pennathur2007,Ren2008a,VanderHeyden2006,Heyden2007,Xie2008,Yang2003}
    Theoretical predictions of the efficiency of standard micro/nano-fluidic channels are 2\% for pure water and 7\% for sodium chloride.~\cite{VanderHeyden2006}
    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10\thinspace--\SI{16}{\micro\metre}.~\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with \SI{6}{\micro\metre} pores.~\cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a \SI{75}{\nano\metre} by \SI{50}{\micro\metre} silica channel.~\cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a \SI{200}{\nano\metre} pore in an alumina membrane.~\cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a \SI{0.5}{\nano\metre} cylindrical pore in polyethylene terephalate foil.~\cite{Xie2008}
    \end{itemize}
    These results indicate that small channels using solutions containing salt are more efficient.
    According to \cite{Daiguji2004b}, the efficiency is maximised when the channel height twice that of the Debye length.
    For water with a sodium content in the range 1--100~ppm this equates to a channel height in the order of 100--10~nm.
    Additionally, ~\cite{VanderHeyden2006} states that the maximum efficiency is found when the salt concentration is low.

@@

    \section{Sizing of Streaming Cell Harvester}
    \label{sect:harvesterSize}
    We now estimate the physical size of a harvester capable of supplying the energy to run a wireless water meter.
    Auckland tap water has a sodium content of 2.1--26.0\thinspace mg/L.~\cite{WatercareNewZealand2012}
    This equates to a Debye length in the range 0.3--\SI{1.0}{\nano\meter} and the optimum channel height between 0.6--\SI{2.1}{\nano\meter}.
    This gap is negligible compared to the thickness of the glass substrate required to create such a harvester, which we presume to be \SI{1}{\milli\meter}.
    The harvester would be made from stacked glass substrates, each separated by a dimension that provides the target 1\% efficiency.
    For simplicity we constrain the dimensions of the channel ends to be square, i.e. the width of the channels will equal the height of the stack.
    The length of the channels in the stack will be \SI{1}{\centi\meter}.
